---
title: "Data607 Neo4J NoSQL"
author: "Yuen Chun Wong"
date: "November 17, 2017"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
      
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(knitr  )
library(getPass)


Neo4J_Username <- getPass::getPass("Please enter your Neo4J username")

Neo4J_Password <- getPass::getPass("Please enter your Neo4J password")







```

### Neo4J Query Lab

## Part I - Load nodes and add relationships

```{r load-csv-file}

library(RCurl)
myfile_flight <- getURL('https://raw.githubusercontent.com/yuenwong2sps/Data607/master/Week12/neo4j-flight-lab-data.csv', 
                 ssl.verifyhost=FALSE, ssl.verifypeer=FALSE)

myfile_airport <- getURL('https://raw.githubusercontent.com/yuenwong2sps/Data607/master/Week12/neo4j-airport-csv-raw.csv', 
                        ssl.verifyhost=FALSE, ssl.verifypeer=FALSE)

mydat_airport <- read.csv(textConnection(myfile_airport), header=T)


mydat_flight <- read.csv(textConnection(myfile_flight), header=T)


```

```{r connect-Neo4J}

library(RNeo4j)




graph = startGraph("http://localhost:7474/db/data/",
                   username = Neo4J_Username,
                   password = Neo4J_Password)

clear(graph, input = FALSE)

```

1. Use Load CSV to create the airport nodes. Note that each airport node should have a label of "Airport" and aset of three attributes: code, city, and state.

```{r load-airport}

#unqiue airport data
addConstraint(graph, "Airport","label")

for(i in 1:length(mydat_airport$label))
{
  createNode(graph, "Airport", 
             label = mydat_airport[i,]$label,
             code=mydat_airport[i,]$label, 
             city=mydat_airport[i,]$city,
             state = mydat_airport[i,]$state
             )
  
}

```

2. Use Load CSV to create the flight nodes. Each flight node should have a label of "Flight" and should have three attributes: number, capacity, and airline.


```{r load-flights}

for(i in 1:length(mydat_flight$flight))
{
  createNode(graph, "flights",
             
             flight = mydat_flight[i,]$flight,
             airline = mydat_flight[i,]$airline,
             capacity = mydat_flight[i,]$capacity
            
             )
  
}

```




3. Use Load CSV to create the arrival relationships. Each arrival relationship should have a label of "Arrive" and should have a single attribute: landing. (This attribute gives the scheduled landing time.)

```{r add-relationship-arrival}
query = "match (a:Airport {code:{airport_code}}), (b:flights {flight:{flight_num}}) create (b) -[r:Arrive {landing:{landing_time}}]->(a)"

t = newTransaction(graph)
for(i in 1:length(mydat_flight$flight))
{
  v_airport_code = mydat_flight[i,]$arrive
  v_flight_num = mydat_flight[i,]$flight
  v_landing_time = mydat_flight[i,]$landing
  
  appendCypher(t,
               query,
               airport_code = v_airport_code,
               flight_num = v_flight_num,
               landing_time = v_landing_time
               )
  
}
commit(t)



```

4. Use Load CSV to create the departure relationships. Each departure relationship should have a label of "Depart" and should have a single attribute: takeoff. (This attribute gives the scheduled takeoff time.)


```{r add-relationship-departure}
query = "match (a:Airport {code:{airport_code}}), (b:flights {flight:{flight_num}}) create (a) -[r:Depart {takeoff:{takeoff_time}}]->(b)"

t = newTransaction(graph)
for(i in 1:length(mydat_flight$flight))
{
  v_airport_code = mydat_flight[i,]$depart
  v_flight_num = mydat_flight[i,]$flight
  v_takeoff_time = mydat_flight[i,]$takeoff
  
  appendCypher(t,
               query,
               airport_code = v_airport_code,
               flight_num = v_flight_num,
               takeoff_time = v_takeoff_time
               )
  
}
commit(t)



```

5. Run the query match (n) return (n)

```{r query-mach_n}
#we need to specific the field to query data instead of just "n"
query = "MATCH (d:Airport)-[t:Depart]->(f:flights)-[l:Arrive]->(a:Airport)
RETURN d.label as FromAirport,t.takeoff as TakeOff, f.airline as AirLine, f.flight as FlightNum,l.landing as Landing, a.label as ToAirport"

match_n_return_n = cypher(graph, query)

kable(match_n_return_n)
```



## Part II: Data Analysis
Now that we have some data loaded, let's answer some basic questions. Write Cypher queries to answer each of the following:

6. Return all flights that run from Detroit to Atlanta.
```{r all-flights-from-Detroit-to-Atlanta}
query = "MATCH (d:Airport)-->(f:flights)-->(a:Airport) where d.code = 'DTW' and a.code = 'ATL'
RETURN d.label as FromAirport, f.airline as AirLine, f.flight as FlightNum, a.label as ToAirport"

flights_From_DTW_To_ATL = cypher(graph, query)

kable(flights_From_DTW_To_ATL)
```

7. Return all flights that take off before 11 a.m.
```{r all-flights-take-off-before-11am}
query = "MATCH (d:Airport)-[t:Depart]->(f:flights)-->(a:Airport) where t.takeoff < 1100
RETURN d.label as FromAirport,t.takeoff as TakeOff, f.airline as AirLine, f.flight as FlightNum, a.label as ToAirport"

flights_Before11AM = cypher(graph, query)

kable(flights_Before11AM)
```

8. Return all flights with a capacity greater than 150 passengers.
```{r all-flights-capacity-larger-than-150}
query = "MATCH (d:Airport)-[t:Depart]->(f:flights)-[l:Arrive]->(a:Airport) where f.capacity > 150
RETURN d.label as FromAirport,t.takeoff as TakeOff, f.airline as AirLine, f.flight as FlightNum,l.landing as Landing, a.label as ToAirport, f.capacity as Capacity"

flights_LargerThan150 = cypher(graph, query)

kable(flights_LargerThan150)
```

9. Return all flights on Delta that arrive in Boston.

```{r all-flights-Delta-arrival-BOS}
query = "MATCH (d:Airport)-[t:Depart]->(f:flights)-[l:Arrive]->(a:Airport) where f.airline ='Delta' and a.code = 'BOS'
RETURN d.label as FromAirport,t.takeoff as TakeOff, f.airline as AirLine, f.flight as FlightNum,l.landing as Landing, a.label as ToAirport"

flights_Delta_BOS = cypher(graph, query)

kable(flights_Delta_BOS)
```


## Part III: Data Modification

Let's make a few simple changes to our data. Perform the following tasks:

10. Suppose the Pittsburgh airport designation code has been changed from PIT to PGH. Write a query to update the airport's code.
```{r update-airport-code}

 Pittsburge_node = getNodes(graph, "Match (a:Airport) where a.code = 'PIT' return a ")

 #updateProp expect a node, but getNodes return a list
 #https://tutel.me/c/programming/questions/35876340/cannot+use+method+39updateprop39+in+rneo4j+package
 Pittsburge_node = Pittsburge_node[[1]] #fix the issue
```

 Before update:
```{r update-airport-code2} 
 (Pittsburge_node)

```

  After update:
```{r update-airport-code3}
 Pittsburge_node = updateProp(Pittsburge_node, code = 'PGH', label= 'PGH')
 
  
 (Pittsburge_node)

```


11. Delta has decided to renumber some of its flights. Write a query to change flight 28 to flight 29.

```{r update-delta-28-to-29}

 delta_node = getNodes(graph, "match (f:flights) where f.airline='Delta' and f.flight=28  return f")

delta_node = delta_node[[1]] #fix the issue

```

 Before udpate:
```{r update-delta-28-to-29-part2}
 (delta_node)
 
```
 
 After update:
```{r update-delta-28-to-29-part3}

 delta_node = updateProp(delta_node, flight = 29)
 

 (delta_node)

```




12. Southwest has cancelled all flights between Detroit and Boston. Write a query that removes all such flights.

```{r delete-Southwest-flights-Detroit-Boston}
query = "match (fa:Airport)-[d:Depart]->(f:flights)-[l:Arrive]->(ta:Airport) where (fa.code = 'DTW' or fa.code = 'BOS') and (ta.code = 'BOS' or ta.code = 'DTW') and f.airline = 'Southwest' delete d, f, l"
 
 cypher(graph, query)

 
 
```

Re-run records to verify the result
```{r query-mach_n2}

query = "MATCH (d:Airport)-[t:Depart]->(f:flights)-[l:Arrive]->(a:Airport)
RETURN d.label as FromAirport,t.takeoff as TakeOff, f.airline as AirLine, f.flight as FlightNum,l.landing as Landing, a.label as ToAirport"

match_n_return_n = cypher(graph, query)

kable(match_n_return_n)
```


## Part IV: Clean Up

Just run cypher to delete all records.
```{r delete-all-records}
query = "match (a) optional match (a)-[r]-() delete a,r"
 
 cypher(graph, query)

 
 
```
