---
title: "Data 607 Week 6 -Project 2"
author: "Yuen Chun Wong"
date: "October 8, 2017"
output: html_document
---

## Tidying and Transforming Data

Load libraries
```{r load-library}
library(tidyr)
library(dplyr)
```



## #1 NYC Foreclosure Data - Sedo Oh
Source: Table 1 (in the pdf file) http://furmancenter.org/files/sotc/SOC_2016_Full.pdf

Analysis: draw tread line of the foreclosure

Load Data - NYC Foreclosure
```{r load-data-NYCForeclosure}
NYCForeclosure <- tbl_df(read.csv("foreclosure.csv", stringsAsFactors = FALSE, check.names = FALSE))
#Boro Summary
print(NYCForeclosure)
```


Tidy the Foreclosure data 
```{r tidy-data-NYCForeclosure}
#get boro

boroType <- NYCForeclosure$Boro

#tidy data
NYCForeclosure.ByYear <- NYCForeclosure %>%
        gather(Yr, ForeclosureCount, 2:7)

print(NYCForeclosure.ByYear)





```

Modifed from http://www.statmethods.net/graphs/line.html
```{r draw-trend-nycforeclosure}
#get the range for x,y
xrange = range(NYCForeclosure.ByYear$Yr)
yrange = range(NYCForeclosure.ByYear$ForeclosureCount)

#setup plot

plot(xrange, yrange, type="n", xlab="Year",
  	ylab="Foreclosure Count" ) 
colors <- rainbow(length(boroType)) 
linetype <- c(1:length(boroType)) 
plotchar <- seq(18,18+length(boroType),1)

# add lines 
for (i in 1:length(boroType)) { 
  boroforeclosure <- subset(NYCForeclosure.ByYear, NYCForeclosure.ByYear$Boro==boroType[i]) 
  lines(boroforeclosure$Yr, boroforeclosure$ForeclosureCount, type="b", lwd=1.5,
    lty=linetype[i], col=colors[i], pch=plotchar[i]) 
} 

# add a title and subtitle 
title("NYC Foreclosure", "")

# add a legend 
legend( "topright", boroType, cex=0.8, col=colors,
  	pch=plotchar, lty=linetype, title="Boro")

```
As we can see above, NYC foreclosure rate gradually decreased but it slightly picked up again from 2015 to 2016.


## #2 Oil supply and demand, price - Silverio Vasquez - EIA Short-term Energy Outlook
Tidy data can be used to downstream analysis.
I saved the data in the csv format and add grouping for easy calcuation before tidying

```{r load-data-oil_suppoly_demand_price}
EIA_OilSupply <- tbl_df(read.csv("EIA_OilSupply.csv", stringsAsFactors = FALSE, check.names = FALSE))
EIA_OilConsumption <- tbl_df(read.csv("EIA_OilConsumption.csv", stringsAsFactors = FALSE, check.names = FALSE))
EIA_OilPrice <- tbl_df(read.csv("EIA_OilPrice.csv", stringsAsFactors = FALSE, check.names = FALSE))

#all oil data
print(EIA_OilSupply)
print(EIA_OilConsumption)
print(EIA_OilPrice)

```


Tidy the Oil data 
```{r tidy-data-Oil-Data}


#tidy data
EIA_OilSupply.ByMonth <- EIA_OilSupply %>%
        gather(Month, MilBarrels, 4:75)

EIA_OilConsumption.ByMonth <- EIA_OilConsumption %>%
        gather(Month, MilBarrels, 4:75)

EIA_OilPrice.ByMonth <- EIA_OilPrice %>%
        gather(DT, PerBarrel, 4:75)


head(EIA_OilSupply.ByMonth)
head(EIA_OilConsumption.ByMonth)
head(EIA_OilPrice.ByMonth)


```

Data above can be merged and summarized.
E.g. US import/consumption vs import avg price.
```{r merged-summarize}
tmp_month <- names(EIA_OilSupply)
Months <- tmp_month[4:75]

US_oil.supply <- EIA_OilSupply.ByMonth %>% filter(DESC == "U.S. (50 States)")
US_oil.Consume_50_States <- EIA_OilConsumption.ByMonth %>% filter(DESC == "U.S. (50 States)")
US_oil.Consume_US_Territories <- EIA_OilConsumption.ByMonth %>% filter(DESC == "U.S. Territories")
US_oil.import <- EIA_OilPrice.ByMonth %>% filter(CODE == "RAIMUUS")

US_oil <- data.frame(
  Record_Months <- Months,
  Supply <- US_oil.supply$MilBarrels,
  Consume <- US_oil.Consume_50_States$MilBarrels + US_oil.Consume_US_Territories$MilBarrels,
  ImportAvgPrice <- US_oil.import$PerBarrel 
  
)

print(US_oil)


```




## #3 NYPD Motor Vehicle Collisions - Yuen Chun Wong
http://www1.nyc.gov/site/nypd/stats/traffic-data/traffic-data-collision.page
It is Aug 2017 data.
I saved the data in csv format and merge the column for location.

```{r load-data-NYPD-Collision}
NYPDCollision <- tbl_df(read.csv("NYPDAug2017_Motor_Vehicle_Collisions.csv", stringsAsFactors = FALSE, check.names = FALSE))
#snapshot of data
head(NYPDCollision)
```


```{r tidy-data-NYPD-collision}

NYPDCollision.ByInjuryType <- NYPDCollision %>% gather(InJury_type, Freq, 4:15)


```


Analysis: Min & Max by Injury Type 
modified from
https://stackoverflow.com/questions/31284794/how-do-i-use-group-by-with-order-function-in-r
https://stackoverflow.com/questions/31528981/select-first-and-last-row-from-grouped-data

```{r max-Injury}
NYPDCollision.ByInjuryType_Max <- NYPDCollision.ByInjuryType %>% group_by(InJury_type) %>% arrange(-Freq) %>%
  filter(row_number()==1)

NYPDCollision.ByInjuryType_Min <- NYPDCollision.ByInjuryType %>% group_by(InJury_type) %>% arrange(Freq) %>%
  filter(row_number()==1)

(NYPDCollision.ByInjuryType_Max)
(NYPDCollision.ByInjuryType_Min)
```
As data above, We can conclude:
1. Queens 109th Precinct has highest collison rate while Manhattan Central Park Precinct has lowerest collison rate.  
2. More Traffic police and road design may be required in Queens 109th Precinct.
3. Few vehicles are Manhattan Central Park Precinct area.

