---
title: "Data607 Project 3"
author: "Yuen Chun Wong"
date: "October 22, 2017"
output: html_document
---


## Project 3 - Most Wanted skills in Finance and Insurance?

```{r load-library}

library(tidyr)
library(dplyr)
library(ggplot2)

```


```{r load-and-prepare-data , warning=FALSE, message=FALSE}


#Clean job data for project 3 across industries
JobData <- tbl_df(read.csv("Project3_CleanData.csv", stringsAsFactors = FALSE, check.names = FALSE))


```

Which skill sets are the most freqents in terms of keywords?
```{r analyse-data}

FinanceAndInsurData_SkillSetCount <- JobData %>%
  filter(Industry == "Finance") %>%
  count(`Skill Set`, sort= TRUE)

# count() is a short-hand for group_by() + tally()

(FinanceAndInsurData_SkillSetCount)

```


As we can see above, Programming/Technical and soft skills are most important while business related keywords are much less frequent.  It may imply that data scientist job in Financial / Insurance job, business knowledge is not the major focus.

sO, WHICH Skill Type are most variable in Programming/Technical and Soft Skills? 
```{r analyssi-data2}

FinanceAndInsurData_Programming_tech <- JobData %>%
  filter(Industry == "Finance" & `Skill Set` == "Programming/Technical") %>%
  count(`Skill Type`, sort=TRUE)


    
(FinanceAndInsurData_Programming_tech)


```

##For programming and technical skills
Python, Java, R, Hadoop are must have. Some skills such as machine learning and cloud technlogy are important too but it could be platform specific (e.g. Google Machine learning)


```{r analysis-data-softskills}

FinanceAndInsurData_SoftSkills <- JobData %>%
  filter(Industry == "Finance" & `Skill Set` == "Soft Skill") %>%
  count(`Skill Type`, sort=TRUE)

(FinanceAndInsurData_SoftSkills)


```

##For softskills
Communication and collaboration are the most important while personality such as Friendly, Positive, Curiosity, Detail are also considered.

```{r draw-softskill}

ggplot(FinanceAndInsurData_SoftSkills, aes(x = FinanceAndInsurData_SoftSkills$`Skill Type`, y = FinanceAndInsurData_SoftSkills$n, fill = FinanceAndInsurData_SoftSkills$n)) + 
  geom_bar(stat = "identity") +
  xlab("Skills") + 
  ylab("Freq.") + 
  theme(legend.position = "none",  
        axis.text.x = element_text(angle = 65, hjust = 1)) +
  ggtitle("Soft Skills of Data Scientist in Finance/Insurance")  
  
```


##Conclusion
So for Finance industry, hard skills such as Python, Java, R, Hadoop are must Have. Be able to communicate and work together are most important soft skills.

