---
title: 'Mushroom Analysis '
output:
  pdf_document: default
  html_document: default
---

## Overview

Smell is very important for cooking, we are going to analyse this mushroom dataset with odor. https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data

## Load Libararies
```{r Load Libraries}
library(devtools)
library(RCurl)
library(bitops)

```

##Load raw data and add labels to the data frame
```{r Load Data}
File_rawMushRoom <- getURL("https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data")
Dataframe_rawMushRoom<-data.frame(read.csv(text=File_rawMushRoom, header=F))
names(Dataframe_rawMushRoom) = c('is_edible','cap_shape',
                                 'cap_surface',
                                 'cap_color',
                                 'IsBruises',
                                 'odor',
                                 'gill_attachment',
                                 'gill_spacing',
                                 'gill_size',
                                 'gill_color',
                                 'stalk_shape',
                                 'stalk_root',
                                 'stalk_surface_above_ring',
                                 'stalk_surface_below_ring',
                                 'stalk_color_above_ring',
                                 'stalk_color_below_ring',
                                 'veil_type',
                                 'veil_color',
                                 'ring_number',
                                 'ring_type',
                                 'spore_print_color',
                                 'population',
                                 'habitat')


```

## Aanlyse data


What is the total count of the mushroom dataset (edible and poisonous)? Given 'e' = edible and 'p' = poisonous.

```{r Get Total mushroom count}

count_mushroom <- length(Dataframe_rawMushRoom$is_edible)

```

How does the edible mushroom smell like?  How many are they?
```{r edible mushroom}
edible_mushroom <- subset(Dataframe_rawMushRoom, Dataframe_rawMushRoom$is_edible == 'e')

odor_edible_mushroom <- table(edible_mushroom$odor)

count_edible_mushroom <- length(edible_mushroom$odor)

```

How does the poisonous mushroom smell like? How many are they?

```{r poisonous mushroom}
poisonous_mushroom <- subset(Dataframe_rawMushRoom, Dataframe_rawMushRoom$is_edible == 'p')

odor_poisonous_mushroom <- table(poisonous_mushroom$odor)

count_poisonous_mushroom <- length(poisonous_mushroom$odor)

```

Columns order are sorted in table, rename columns correspondingly

```{r format table results}

new_odor_labels = c('almond','creosote','foul','anise',
                    'musty','none', 'pungent','fishy','spicy')

names(odor_edible_mushroom ) = new_odor_labels
names(odor_poisonous_mushroom ) = new_odor_labels


```

##Result

Couts:
```{r Total mushroom count}

paste("Total Mushroom in our dataset:", as.character( count_mushroom))
paste("Total edible:", as.character( count_edible_mushroom))
paste("Total poison:", as.character(count_poisonous_mushroom))

```

Odor for edible mushrooms 
```{r Output edible mushroom}
print(odor_edible_mushroom)
barplot(odor_edible_mushroom, main="Odor for Edible Mushrooms")
```

And odor for poison mushrooms
```{r Output poison mushroom}
print(odor_poisonous_mushroom)
barplot(odor_poisonous_mushroom, main="Odor for Poisonous Mushrooms")
```


##Observation:

1) the mushroom with almond or anise are edible
2) the mushroom with creosote, foul, musty, pungent, spicy and fishy are poison
3) the mushroom without odor mostly edible (3408)/(3408+120) = 96.6%
4) it seems that mushroom with 'bad' smell mostly are poison.  Edbile mushroom is not only source of food and it could enhance the smell of the dish.



