---
title: "Data 607 Week 3"
author: "Yuen Chun Wong"
date: "September 17, 2017"
output: pdf_document
---

##Week 3



Given a string 
```{r load-data}

library(stringr)

raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"  

print( raw.data )

```

3.1) rearrange the vector so that all element conform to the standard first_name lastname

```{r ans1}
raw.data.fullname <- unlist(str_extract_all(raw.data, "[[:alpha:][:space:].,]{2,}"))
raw.data.fullname

raw.data.lastname <- unlist(str_extract_all(raw.data.fullname, "[[:alpha:]]{2,},"))

raw.data.lastname

raw.data.lastname2 <- unlist(str_extract_all(str_extract_all(raw.data.fullname, "[^,][:alpha:][:space:][:alpha:]{2,}"),"[:space:][:alpha:]{2,}"))
raw.data.lastname2


raw.data.firstname <- unlist(str_extract_all(raw.data.fullname, "[,][.[:space:][:alpha:]]{2,}"))

raw.data.firstname

raw.data.firstname2 <- unlist(str_extract_all(raw.data.fullname, "[^,][:alpha:]{2,}[:space:]"))
raw.data.firstname2

raw.data.firstname3 <- c(raw.data.firstname, raw.data.firstname2)
raw.data.lastname3 <- c(raw.data.lastname, raw.data.lastname2)

#use str_replace_all()
raw.data.firstname3 <- str_replace_all(raw.data.firstname3,",","")
raw.data.lastname3 <- str_replace_all(raw.data.lastname3,",","")

df <- data.frame(
      firstname = raw.data.firstname3,
      lastname = raw.data.lastname3
                 )

print(df)
```


3.2) Construct a logical vector indicating whether a character has a title (ie. Rev. And Dr)

Try to construct regular expression to detect title which beginning with no space, contain a-z, A-Z and follow by "."
```{r detect_title1}
title_regex <- "[[:alpha:]]{2,}[.]"
WithTitle <- unlist(str_extract_all(raw.data.fullname,title_regex))
print(WithTitle)

```

Build a table to show the result
```{r detect_title2}
chk_title <- data.frame(
        fullname <- unlist(raw.data.fullname),
        HasTitle <- str_detect(raw.data.fullname, title_regex)
        )

print(chk_title)

```



3.3) Construct a logical vector indicating whether a character has a second name

construct a logic to detech second name that has "space" in the name between "words"
```{r detect_secondname1}
secondname_regex <- "[[:alpha:].]{2,}[[:space:]][[:alpha:]]{2,}"
WithsecondName <- unlist(str_extract_all(raw.data.firstname3,secondname_regex))
print(WithsecondName)

```

Build a table to show the result
```{r detect_secondname2}
chk_secondname <- data.frame(
        firstname <- unlist(raw.data.firstname3),
        HasTitle <- str_detect(raw.data.firstname3, secondname_regex)
        )

print(chk_secondname)

```



4)Describe the types of strings that conform to the following regular expressions and
construct an example that is matched by the regular expression.
(1) [0-9]+\\$
 Ans: match all numberic number, at least one digit, and then end of the line.  "[[:digit:]]+"

(2) \\b[a-z]{1,4}\\b
  Ans: with word bounds at the begining and the end of the word, word with lower case alphabetic, at least one characters, max 4 characters.
  "\\b[[:lower:]]{1,4}\\b"
  
(3) .*?\\.txt$
  Ans: match anything from zero to many times follow by ".txt"
   "[[:alnum:]]*.txt"

(4) \\d{2}/\\d{2}/\\d{4}
   
  Ans: It is the date format   mm/dd/yyyy format.  
  "[[:digit:]]{2}/[[:digit:]]{2}/[[:digit:]]{4}""

(5) <(.+?)>.+?</\\1>
  Ans:  <match anyting>match anyting</first match reference>
  "<[[:alnum:]]{1,}>[[:alnum:]]{1,}</\\1>"
